---
title: "cosc421_project"
author: "Angelina Dubrule"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=3)
options(repos = c(CRAN = "https://cran.rstudio.com/"))
install.packages("igraph")
library("igraph")
```

## Generate Network Graph Whole Network

```{r graph}
nodes <- read.csv("Data_Nodes.csv", header = T)
edges <- read.csv("Data_Edges.csv", header = T)

graph <- graph_from_data_frame(edges, nodes, directed = T)

plot(
  graph,
  vertex.label = NA,
  vertex.size = 5,
  main = "Mobi Vancouver Bike Share Network",
  edge.arrow.size = 0.5
)
```

## In-Degree Whole Network

```{r indeg}
indeg <- degree(graph, mode = "in")
cat("In-Degree Distribution:", indeg, "\n\n")

topfive <- order(indeg, decreasing = TRUE)[1:5]
topnodes <- names(indeg)[topfive]
topvalues <- indeg[topfive]

cat("Top 5 Nodes by In-Degree:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with In-Degree:", topvalues[i], "\n")
}

hist(
  indeg,
  breaks = seq(min(indeg) - 0.5, max(indeg) + 0.5, by = 1),
  col = "lightblue",
  border = "black",
  probability = T,
  main = "Bike Share In-Degree Distribution",
  xlab = "In-Degree",
  ylab = "Density",
  ylim = c(0, 0.2)
)

curve(dnorm(x, mean = mean(indeg), sd = sd(indeg)), col = "darkblue", lwd = 2, add = T)
```

## Out-Degree Whole Network

```{r outdeg}
outdeg <- degree(graph, mode = "out")
cat("Out-Degree Distribution:", outdeg, "\n\n")

topfive <- order(outdeg, decreasing = TRUE)[1:5]
topnodes <- names(outdeg)[topfive]
topvalues <- outdeg[topfive]

cat("Top 5 Nodes by Out-Degree:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with Out-Degree:", topvalues[i], "\n")
}

sinks <- names(outdeg)[outdeg == 0 && indeg != 0]
cat("\nSink Nodes (Out-Degree = 0):\n")
if (length(sinks) > 0) {
  cat(sinks, sep = ", ")
} else {
  cat("No sink nodes found.")
}

hist(
  outdeg,
  breaks = seq(min(outdeg) - 0.5, max(outdeg) + 0.5, by = 1),
  col = "lightgreen",
  border = "black",
  probability = T,
  main = "Bike Share Out-Degree Distribution",
  xlab = "Out-Degree",
  ylab = "Density",
  ylim = c(0, 0.2)
)

curve(dnorm(x, mean = mean(outdeg), sd = sd(outdeg)), col = "darkgreen", lwd = 2, add = T)
```

## Average Degree Whole Network

```{r avgdeg}
avgdeg <- mean(degree(graph))
cat("Average Degree:", avgdeg, "\n\n")
cat("Average In/Out-Degree:", avgdeg / 2)
```

## Degree Ratio Whole Network

```{r degratio}
degratio <- indeg / outdeg

topfive <- order(degratio, decreasing = TRUE)[1:5]
topnodes <- names(degratio)[topfive]
topvalues <- degratio[topfive]

cat("Top 5 Nodes with Highest Degree Ratio:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with Degree Ratio:", topvalues[i], "\n")
}

topfive <- order(degratio, decreasing = FALSE)[1:5]
topnodes <- names(degratio)[topfive]
topvalues <- degratio[topfive]

cat("\nTop 5 Nodes with Lowest Degree Ratio:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with Degree Ratio:", topvalues[i], "\n")
}

```

## Generate Network Cluster 1

```{r graph1}
nodes <- read.csv("Cluster_1_Nodes.csv", header = T)
edges <- read.csv("Cluster_1_Edges.csv", header = T)
graph <- graph_from_data_frame(edges, nodes, directed = T)
```

## In-Degree Cluster 1

```{r indeg1}
indeg <- degree(graph, mode = "in")
cat("In-Degree Distribution:", indeg, "\n\n")

topfive <- order(indeg, decreasing = TRUE)[1:5]
topnodes <- names(indeg)[topfive]
topvalues <- indeg[topfive]

cat("Top 5 Nodes by In-Degree:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with In-Degree:", topvalues[i], "\n")
}

hist(
  indeg,
  breaks = seq(min(indeg) - 0.5, max(indeg) + 0.5, by = 1),
  col = "lightblue",
  border = "black",
  probability = T,
  main = "Bike Share In-Degree Distribution",
  xlab = "In-Degree",
  ylab = "Density",
  ylim = c(0, 0.2)
)

curve(dnorm(x, mean = mean(indeg), sd = sd(indeg)), col = "darkblue", lwd = 2, add = T)
```

## Out-Degree Cluster 1

```{r outdeg1}
outdeg <- degree(graph, mode = "out")
cat("Out-Degree Distribution:", outdeg, "\n\n")

topfive <- order(outdeg, decreasing = TRUE)[1:5]
topnodes <- names(outdeg)[topfive]
topvalues <- outdeg[topfive]

cat("Top 5 Nodes by Out-Degree:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with Out-Degree:", topvalues[i], "\n")
}

sinks <- names(outdeg)[outdeg == 0 && indeg != 0]
cat("\nSink Nodes (Out-Degree = 0):\n")
if (length(sinks) > 0) {
  cat(sinks, sep = ", ")
} else {
  cat("No sink nodes found.")
}

hist(
  outdeg,
  breaks = seq(min(outdeg) - 0.5, max(outdeg) + 0.5, by = 1),
  col = "lightgreen",
  border = "black",
  probability = T,
  main = "Bike Share Out-Degree Distribution",
  xlab = "Out-Degree",
  ylab = "Density",
  ylim = c(0, 0.2)
)

curve(dnorm(x, mean = mean(outdeg), sd = sd(outdeg)), col = "darkgreen", lwd = 2, add = T)
```

## Average Degree Cluster 1

```{r avgdeg1}
avgdeg <- mean(degree(graph))
cat("Average Degree:", avgdeg, "\n\n")
cat("Average In/Out-Degree:", avgdeg / 2)
```

## Degree Ratio Cluster 1

```{r degratio1}
degratio <- indeg / outdeg

topfive <- order(degratio, decreasing = TRUE)[1:5]
topnodes <- names(degratio)[topfive]
topvalues <- degratio[topfive]

cat("Top 5 Nodes with Highest Degree Ratio:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with Degree Ratio:", topvalues[i], "\n")
}

topfive <- order(degratio, decreasing = FALSE)[1:5]
topnodes <- names(degratio)[topfive]
topvalues <- degratio[topfive]

cat("\nTop 5 Nodes with Lowest Degree Ratio:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with Degree Ratio:", topvalues[i], "\n")
}

```

## Generate Network Cluster 2

```{r graph2}
nodes <- read.csv("Cluster_2_Nodes.csv", header = T)
edges <- read.csv("Cluster_2_Edges.csv", header = T)

graph <- graph_from_data_frame(edges, nodes, directed = T)

plot(
  graph,
  vertex.label = NA,
  vertex.size = 5,
  main = "Mobi Vancouver Bike Share Network",
  edge.arrow.size = 0.5
)
```

## In-Degree Cluster 2

```{r indeg2}
indeg <- degree(graph, mode = "in")
cat("In-Degree Distribution:", indeg, "\n\n")

topfive <- order(indeg, decreasing = TRUE)[1:5]
topnodes <- names(indeg)[topfive]
topvalues <- indeg[topfive]

cat("Top 5 Nodes by In-Degree:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with In-Degree:", topvalues[i], "\n")
}

hist(
  indeg,
  breaks = seq(min(indeg) - 0.5, max(indeg) + 0.5, by = 1),
  col = "lightblue",
  border = "black",
  probability = T,
  main = "Bike Share In-Degree Distribution",
  xlab = "In-Degree",
  ylab = "Density",
  ylim = c(0, 0.2)
)

curve(dnorm(x, mean = mean(indeg), sd = sd(indeg)), col = "darkblue", lwd = 2, add = T)
```

## Out-Degree Cluster 2

```{r outdeg2}
outdeg <- degree(graph, mode = "out")
cat("Out-Degree Distribution:", outdeg, "\n\n")

topfive <- order(outdeg, decreasing = TRUE)[1:5]
topnodes <- names(outdeg)[topfive]
topvalues <- outdeg[topfive]

cat("Top 5 Nodes by Out-Degree:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with Out-Degree:", topvalues[i], "\n")
}

sinks <- names(outdeg)[outdeg == 0 && indeg != 0]
cat("\nSink Nodes (Out-Degree = 0):\n")
if (length(sinks) > 0) {
  cat(sinks, sep = ", ")
} else {
  cat("No sink nodes found.")
}

hist(
  outdeg,
  breaks = seq(min(outdeg) - 0.5, max(outdeg) + 0.5, by = 1),
  col = "lightgreen",
  border = "black",
  probability = T,
  main = "Bike Share Out-Degree Distribution",
  xlab = "Out-Degree",
  ylab = "Density",
  ylim = c(0, 0.2)
)

curve(dnorm(x, mean = mean(outdeg), sd = sd(outdeg)), col = "darkgreen", lwd = 2, add = T)
```

## Average Degree Cluster 2

```{r avgdeg2}
avgdeg <- mean(degree(graph))
cat("Average Degree:", avgdeg, "\n\n")
cat("Average In/Out-Degree:", avgdeg / 2)
```

## Degree Ratio Cluster 2

```{r degratio2}
degratio <- indeg / outdeg

topfive <- order(degratio, decreasing = TRUE)[1:5]
topnodes <- names(degratio)[topfive]
topvalues <- degratio[topfive]

cat("Top 5 Nodes with Highest Degree Ratio:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with Degree Ratio:", topvalues[i], "\n")
}

topfive <- order(degratio, decreasing = FALSE)[1:5]
topnodes <- names(degratio)[topfive]
topvalues <- degratio[topfive]

cat("\nTop 5 Nodes with Lowest Degree Ratio:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with Degree Ratio:", topvalues[i], "\n")
}

```

## Generate Network Cluster 3

```{r graph3}
nodes <- read.csv("Cluster_3_Nodes.csv", header = T)
edges <- read.csv("Cluster_3_Edges.csv", header = T)

graph <- graph_from_data_frame(edges, nodes, directed = T)

plot(
  graph,
  vertex.label = NA,
  vertex.size = 5,
  main = "Mobi Vancouver Bike Share Network",
  edge.arrow.size = 0.5
)
```

## In-Degree Cluster 3

```{r indeg3}
indeg <- degree(graph, mode = "in")
cat("In-Degree Distribution:", indeg, "\n\n")

topfive <- order(indeg, decreasing = TRUE)[1:5]
topnodes <- names(indeg)[topfive]
topvalues <- indeg[topfive]

cat("Top 5 Nodes by In-Degree:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with In-Degree:", topvalues[i], "\n")
}

hist(
  indeg,
  breaks = seq(min(indeg) - 0.5, max(indeg) + 0.5, by = 1),
  col = "lightblue",
  border = "black",
  probability = T,
  main = "Bike Share In-Degree Distribution",
  xlab = "In-Degree",
  ylab = "Density",
  ylim = c(0, 0.2)
)

curve(dnorm(x, mean = mean(indeg), sd = sd(indeg)), col = "darkblue", lwd = 2, add = T)
```

## Out-Degree Cluster 3

```{r outdeg3}
outdeg <- degree(graph, mode = "out")
cat("Out-Degree Distribution:", outdeg, "\n\n")

topfive <- order(outdeg, decreasing = TRUE)[1:5]
topnodes <- names(outdeg)[topfive]
topvalues <- outdeg[topfive]

cat("Top 5 Nodes by Out-Degree:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with Out-Degree:", topvalues[i], "\n")
}

sinks <- names(outdeg)[outdeg == 0 && indeg != 0]
cat("\nSink Nodes (Out-Degree = 0):\n")
if (length(sinks) > 0) {
  cat(sinks, sep = ", ")
} else {
  cat("No sink nodes found.")
}

hist(
  outdeg,
  breaks = seq(min(outdeg) - 0.5, max(outdeg) + 0.5, by = 1),
  col = "lightgreen",
  border = "black",
  probability = T,
  main = "Bike Share Out-Degree Distribution",
  xlab = "Out-Degree",
  ylab = "Density",
  ylim = c(0, 0.2)
)

curve(dnorm(x, mean = mean(outdeg), sd = sd(outdeg)), col = "darkgreen", lwd = 2, add = T)
```

## Average Degree Cluster 3

```{r avgdeg3}
avgdeg <- mean(degree(graph))
cat("Average Degree:", avgdeg, "\n\n")
cat("Average In/Out-Degree:", avgdeg / 2)
```

## Degree Ratio Cluster 3

```{r degratio3}
degratio <- indeg / outdeg

topfive <- order(degratio, decreasing = TRUE)[1:5]
topnodes <- names(degratio)[topfive]
topvalues <- degratio[topfive]

cat("Top 5 Nodes with Highest Degree Ratio:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with Degree Ratio:", topvalues[i], "\n")
}

topfive <- order(degratio, decreasing = FALSE)[1:5]
topnodes <- names(degratio)[topfive]
topvalues <- degratio[topfive]

cat("\nTop 5 Nodes with Lowest Degree Ratio:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with Degree Ratio:", topvalues[i], "\n")
}

```

## Generate Network Cluster 4

```{r graph4}
nodes <- read.csv("Cluster_4_Nodes.csv", header = T)
edges <- read.csv("Cluster_4_Edges.csv", header = T)

graph <- graph_from_data_frame(edges, nodes, directed = T)

plot(
  graph,
  vertex.label = NA,
  vertex.size = 5,
  main = "Mobi Vancouver Bike Share Network",
  edge.arrow.size = 0.5
)
```

## In-Degree Cluster 4

```{r indeg4}
indeg <- degree(graph, mode = "in")
cat("In-Degree Distribution:", indeg, "\n\n")

topfive <- order(indeg, decreasing = TRUE)[1:5]
topnodes <- names(indeg)[topfive]
topvalues <- indeg[topfive]

cat("Top 5 Nodes by In-Degree:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with In-Degree:", topvalues[i], "\n")
}

hist(
  indeg,
  breaks = seq(min(indeg) - 0.5, max(indeg) + 0.5, by = 1),
  col = "lightblue",
  border = "black",
  probability = T,
  main = "Bike Share In-Degree Distribution",
  xlab = "In-Degree",
  ylab = "Density",
  ylim = c(0, 0.2)
)

curve(dnorm(x, mean = mean(indeg), sd = sd(indeg)), col = "darkblue", lwd = 2, add = T)
```

## Out-Degree Cluster 4

```{r outdeg4}
outdeg <- degree(graph, mode = "out")
cat("Out-Degree Distribution:", outdeg, "\n\n")

topfive <- order(outdeg, decreasing = TRUE)[1:5]
topnodes <- names(outdeg)[topfive]
topvalues <- outdeg[topfive]

cat("Top 5 Nodes by Out-Degree:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with Out-Degree:", topvalues[i], "\n")
}

sinks <- names(outdeg)[outdeg == 0 && indeg != 0]
cat("\nSink Nodes (Out-Degree = 0):\n")
if (length(sinks) > 0) {
  cat(sinks, sep = ", ")
} else {
  cat("No sink nodes found.")
}

hist(
  outdeg,
  breaks = seq(min(outdeg) - 0.5, max(outdeg) + 0.5, by = 1),
  col = "lightgreen",
  border = "black",
  probability = T,
  main = "Bike Share Out-Degree Distribution",
  xlab = "Out-Degree",
  ylab = "Density",
  ylim = c(0, 0.2)
)

curve(dnorm(x, mean = mean(outdeg), sd = sd(outdeg)), col = "darkgreen", lwd = 2, add = T)
```

## Average Degree Cluster 4

```{r avgdeg4}
avgdeg <- mean(degree(graph))
cat("Average Degree:", avgdeg, "\n\n")
cat("Average In/Out-Degree:", avgdeg / 2)
```

## Degree Ratio Cluster 4

```{r degratio4}
degratio <- indeg / outdeg

topfive <- order(degratio, decreasing = TRUE)[1:5]
topnodes <- names(degratio)[topfive]
topvalues <- degratio[topfive]

cat("Top 5 Nodes with Highest Degree Ratio:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with Degree Ratio:", topvalues[i], "\n")
}

topfive <- order(degratio, decreasing = FALSE)[1:5]
topnodes <- names(degratio)[topfive]
topvalues <- degratio[topfive]

cat("\nTop 5 Nodes with Lowest Degree Ratio:\n")
for (i in 1:5) {
  cat("Node:", topnodes[i], " with Degree Ratio:", topvalues[i], "\n")
}

```
